Short and Mechanized Logical Relation for Dependent Type Theory

### 2024-02-25 paper/paper.tex
some subsections to organize related work
Extend the feature matrix with Zombie

% \begin{minipage}{0.8\textwidth}
  \begin{tabular}{ l |  l  | l | l | l }
      & \header{Universes} & \header{Inductives} & \header{Convertibility}
      & \header{Large Elim} \\
    \hline
    \lang{} & $\mathbb{N}$ & Id, Bool & U & {\boxedsymbols ✓} \\
    \hline
    NBE-in-Coq & 1 & Nat & T & {\boxedsymbols ✗} \\
    \hline
    POPLmark2 & 1 & Nat & T & {\boxedsymbols ✓} \\
    \hline
    MLTT-a-la-Coq & 1 & Id, Nat & T & {\boxedsymbols ✓} \\
    \hline
    Nuprl & $\mathbb{N}$ & W-Types & E & {\boxedsymbols ✓} \\
    \hline
    Coq-in-Coq & $\mathbb{P}$,$\mathbb{N}$ & General & U & {\boxedsymbols ✓}
    %   & \lang{} & \citet{anand2014towards} &
    %                                          \citet{barras2012semantical}
    % & \citet{martin-lof-a-la-coq}  \\
    % \hline
    % Universe & Countable & Countable & Impredicative + Countable & One\\
    % % \hline
    % % Large Elimination & {\boxedsymbols ✓} & {\boxedsymbols ✓} & {\boxedsymbols ✓} \\
    % \hline
    % Inductive Types & Id, Bool & W types & General inductive types & Nat \\
    % \hline
    % Convertibility & Untyped & Untyped,Extensional & Untyped & Typed \\
  \end{tabular}
% \end{minipage}

  \begin{tabular}{l}
  \\
    NBE-in-Coq \citet{nbeincoq} \\
    POPLmark reloaded \citet{decagda} \\
    MLTT-a-la-Coq \citet{martin-lof-a-la-coq} \\
    Nuprl \citet{anand2014towards} \\
  \\ 
  Universes: Countable ($\mathbb{N}$), One (1), Impredicative ($\mathbb{P}$) \\
  Convertibility: Typed (T), Untyped (U), Extensional (E) \\
  \end{tabular}

  \caption{Overview of Language Features}
  \label{fig:linecount}
\end{figure}

@@

Proof by logical relations does not come with a precise definition. In
the most general sense, a logical relation can be viewed as a
practical technique that uses a type-indexed relation to assign
meanings to types and strengthen the induction hypothesis for the
property of interest. The original idea of this technique can be
traced back to
\citet{tait1967:reducibility}. \citet{tait1967:reducibility} maps
types to sets of terms satisfying certain properties related to reduction.
The same idea is explained in \citet{girard1989proofs} and extended to
prove the strong normalization property of System F.

@@

@@

>>>>

% \begin{minipage}{0.8\textwidth}
  \begin{tabular}{ l |  l  | l | l | l }
      & \header{Universes} & \header{Inductives} & \header{Convertibility}
      & \header{Large Elim} \\
    \hline
    \lang{} & $\mathbb{N}$ & Id, Bool & U & {\boxedsymbols ✓} \\
    % \hline
    $\lambda^\theta$ & 0 & Id, Nat & U & {\boxedsymbols ✗}\\
    NBE-in-Coq & 1 & Nat & T & {\boxedsymbols ✗} \\
    % \hline
    $\lambda^{\Pi U\mathbb{N}}$ & 1 & Nat & T & {\boxedsymbols ✓} \\
    % \hline
    MLTT-a-la-Coq & 1 & Id, Nat & T & {\boxedsymbols ✓} \\
    % \hline
    Core Nuprl & $\mathbb{N}$ & W-Types & E & {\boxedsymbols ✓} \\
    % \hline
    CIC & $\mathbb{I}$,$\mathbb{N}$ & General & U & {\boxedsymbols ✓}\\
    %   & \lang{} & \citet{anand2014towards} &
    %                                          \citet{barras2012semantical}
    % & \citet{martin-lof-a-la-coq}  \\
    % \hline
    % Universe & Countable & Countable & Impredicative + Countable & One\\
    % % \hline
    % % Large Elimination & {\boxedsymbols ✓} & {\boxedsymbols ✓} & {\boxedsymbols ✓} \\
    % \hline
    % Inductive Types & Id, Bool & W types & General inductive types & Nat \\
    % \hline
    % Convertibility & Untyped & Untyped,Extensional & Untyped & Typed \\
  \end{tabular}
% \end{minipage}

  \begin{tabular}{l}
  \\
    $\lambda^\theta$ (logical fragment only) \citet{casinghino:combining-proofs-programs} \\
    NBE-in-Coq \citet{nbeincoq} \\
    $\lambda^{\Pi U\mathbb{N}}$ \citet{decagda} \\
    MLTT-a-la-Coq \citet{martin-lof-a-la-coq} \\
    Core Nuprl \citet{anand2014towards} \\
    CIC \citet{barras2012semantical} \\
  \\
  Universes: Countable ($\mathbb{N}$), Zero (0), One (1), Impredicative ($\mathbb{I}$) \\
  Convertibility: Typed (T), Untyped (U), Extensional (E) \\
  \end{tabular}

  \caption{Feature matrix for dependently typed languages with
    mechanized proofs by logical relation}
  \label{fig:linecount}
\end{figure}

@@

\subsection{Logical relations}
Proof by logical relations does not come with a precise definition. In
the most general sense, a logical relation can be viewed as a
practical technique that uses a type-indexed relation to assign
meanings to types and strengthen the induction hypothesis for the
property of interest. The original idea of this technique can be
traced back to
\citet{tait1967:reducibility}. \citet{tait1967:reducibility} maps
types to sets of terms satisfying certain properties related to reduction.
The same idea is explained in \citet{girard1989proofs} and extended to
prove the strong normalization property of System F.

@@

\subsection{Mechanized logical relations about dependent types}

@@

\subsection{Other mechanized metatheory of dependent types}
